---
title: "Run dmc"
author: "Sarah P. Flanagan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
---

First I'll set up the working environment.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir='../fwsw_results/')
```

```{r source, include=FALSE}
source("../../gwscaR/R/gwscaR.R")
source("../../gwscaR/R/gwscaR_plot.R")
source("../../gwscaR/R/gwscaR_utility.R")
source("../../gwscaR/R/gwscaR_fsts.R")
source("../../gwscaR/R/gwscaR_popgen.R")
library(knitr)
library(fields)
library(MASS)
library(RColorBrewer)
```

and load the necessary files
```{r access_analysis, echo=FALSE}
knitr::read_chunk("../scripts/002_fwsw_analysis.R")
knitr::read_chunk("../scripts/002_run_dmc.R")
```
```{r FWSWsetup, include=FALSE}
```
```{r read_vcf}
vcf<-parse.vcf("p4.upd.vcf") #this is the smaller dataset
```
```{r vcfSetup, echo=TRUE}
```

Now I'll go through the steps outline in https://github.com/kristinmlee/dmc/blob/master/dmc_example.md

## Calculate neutral F matrix

For this first step (calculating the neutral variance/covariance matrix, F), I'll use linkage groups that have no or few shared outlier loci. I need to (1) Calculate allele frequencies for each population, (2) specify a vector of sample sizes for each population, and (3) specify a string for filename for output. This step is baseline for all parameter sensitivity analyses.

```{r calc_af, eval=FALSE}
calc.allFreqs<-function(vcf,pop.list, pop.labs){
  allFreqs<-do.call(rbind,lapply(pop.list, function(pop){
    this.vcf<-cbind(vcf[,1:9],vcf$SNP,vcf[,grep(pop,colnames(vcf))])
    afs<-do.call(rbind,apply(this.vcf,1,calc.afs.vcf))
    return(afs$RefFreq)
  }))
  colnames(allFreqs)<-vcf$SNP
  rownames(allFreqs)<-pop.labs
  return(allFreqs)
}
#calculate selected allele frequencies
LG4<-calc.allFreqs(vcf[vcf$`#CHROM` == "LG4" & vcf$POS < 120000,],pop.list,pop.labs)
saveRDS(LG4,"dmc/selectedRegionAlleleFreqs_p4LG4.RDS")
saveRDS(vcf[vcf$`#CHROM`=="LG4","POS"],"dmc/selectedRegionPositions_p4LG4.RDS")

#calculate neutral allele frequencies
allFreqs<-calc.allFreqs(vcf[vcf$`#CHROM` %in%
                              c("LG3","LG5","LG6","LG7","LG8","LG9","LG11","LG15","LG16","LG17","LG19","LG21","LG22"),],
                   pop.list,pop.labs)
saveRDS(allFreqs,"dmc/neutralAlleleFreqs_p4LG4.RDS")
```
```{r read_allFreqs}
allFreqs<-readRDS("dmc/neutralAlleleFreqs_p4LG4.RDS")
```

Once those are calculated, I can calculate the neutral F matrix, in addition to its determinant and inverse.

```{r calc_neutralF, eval=FALSE}
sampleSizes<-unlist(lapply(pop.list,function(pop){
  n<-length(grep(pop,colnames(vcf)))
  return(2*n) }))
numPops<-16

#calculate the neutral F matrix
saveRDS(sampleSizes,"dmc/sampleSizes.RDS")
neutralF_filename<-"dmc/neutralF_p4LG4_50" #for 50 positions
source("../programs/dmc-master/calcNeutralF.R")
F_estimate<-readRDS("dmc/neutralF_p4LG4_50.RDS")

#calculate the determinant and inverse of the neutral F matrix
M <- numPops
Tmatrix <- matrix(data = rep(-1 / M, (M - 1) * M), nrow = M - 1, ncol = M)
diag(Tmatrix) = (M - 1) / M 
sampleErrorMatrix = diag(1/sampleSizes, nrow = numPops, ncol = numPops)
det_FOmegas_neutral = det(Tmatrix %*% (F_estimate + sampleErrorMatrix) %*% t(Tmatrix))
saveRDS(det_FOmegas_neutral, "dmc/det_FOmegas_neutral_p4LG4_50.RDS")

inv_FOmegas_neutral = ginv(Tmatrix %*% (F_estimate + sampleErrorMatrix) %*% t(Tmatrix))
saveRDS(inv_FOmegas_neutral, "dmc/inv_FOmegas_neutral_p4LG4_50.RDS")

```

```{r calc_neutralF_100, eval=FALSE}
selSite = positions[seq(1, length(positions[positions<12000000]), length.out = 100)]
allFreqs<-readRDS("dmc/neutralAlleleFreqs_p4LG4.RDS")
#calculate the neutral F matrix
sampleSizes<-readRDS("dmc/sampleSizes.RDS")
neutralF_filename<-"dmc/neutralF_p4LG4_100" #for 50 positions
source("../programs/dmc-master/calcNeutralF.R")
F_estimate<-readRDS("dmc/neutralF_p4LG4_100.RDS")

#calculate the determinant and inverse of the neutral F matrix
M <- numPops
Tmatrix <- matrix(data = rep(-1 / M, (M - 1) * M), nrow = M - 1, ncol = M)
diag(Tmatrix) = (M - 1) / M 
sampleErrorMatrix = diag(1/sampleSizes, nrow = numPops, ncol = numPops)
det_FOmegas_neutral = det(Tmatrix %*% (F_estimate + sampleErrorMatrix) %*% t(Tmatrix))
saveRDS(det_FOmegas_neutral, "dmc/det_FOmegas_neutral_p4LG4_100.RDS")

inv_FOmegas_neutral = ginv(Tmatrix %*% (F_estimate + sampleErrorMatrix) %*% t(Tmatrix))
saveRDS(inv_FOmegas_neutral, "dmc/inv_FOmegas_neutral_p4LG4_100.RDS")

```

*Note: I had to use `r Sys.chmod('./dmc/')` to allow read/write in the scripts.*

Let's look at the allele frequencies on LG4. This will help inform the hypotheses I test with dmc.


```{r sig_alleleFreqs}
stacks.sig<-read.delim("p4.stacks.sig.snps.txt")
stacks.sig$SNP<-paste(stacks.sig$Chr,(stacks.sig$BP+1),sep=".")
LG4<-readRDS("dmc/selectedRegionAlleleFreqs_p4LG4.RDS")
allFreqs<-readRDS("dmc/neutralAlleleFreqs_p4LG4.RDS")
# LG4[rownames(LG4) %in% c("FLFW","ALFW","LAFW","TXFW"),colnames(LG4) %in% stacks.sig$SNP]
# LG4[rownames(LG4) %in% c("FLFW","ALFW","LAFW","TXFW"),
#          which(colnames(LG4) %in% stacks.sig$SNP)[1]:
#            which(colnames(LG4) %in% stacks.sig$SNP)[length(which(colnames(LG4) %in% stacks.sig$SNP))]]
grp.colors<-c('#762a83','#af8dc3','#e7d4e8','#d9f0d3','#7fbf7b','#1b7837')
plot(LG4["TXFW",],col=grp.colors[1],pch="-",bty="L",lwd=2,cex=2)
points(LG4["LAFW",],col=grp.colors[2],pch="-",bty="L",lwd=2,cex=2)
points(LG4["ALFW",],col=grp.colors[3],pch="-",bty="L",lwd=2,cex=2)
points(LG4["FLFW",],col=grp.colors[6],pch="-",bty="L",lwd=2,cex=2)

cols<-colorRampPalette(t(brewer.pal(9,name="Blues")[3:7]))(100)

fields::image.plot(t(allFreqs[rownames(allFreqs) %in% c("FLFW","ALFW","LAFW","TXFW"),]),col=cols,
      axes=FALSE)
text(x=-0.02,y=1,"FLFW",srt=90,xpd=TRUE)
text(x=-0.02,y=0.67,"ALFW",srt=90,xpd=TRUE)
text(x=-0.02,y=0.35,"LAFW",srt=90,xpd=TRUE)
text(x=-0.02,y=0,"TXFW",srt=90,xpd=TRUE)
mtext("Position on LG4",1)
axis(3,at =t(LG4[rownames(LG4) %in% c("FLFW","ALFW","LAFW","TXFW"),colnames(LG4) %in% stacks.sig$SNP]),
     labels = FALSE,lwd=0,lwd.ticks = 1.75,tck=-.05)
```

This shows that the FLFW site has somewhat different allele frequencies than the other sites. This suggests that I should test to see whether there was an independent mutation in the FLFW population and the other three have a single sweep.

## Calculate F<sup>(S)</sup> matrices

We generate matrices for the following five models:
*1. All selected pops have independent mutations of beneficial allele
*2. All selected pops share beneficial allele via migration
*3. Beneficial allele was standing the ancestor of all selected pops
*4. There was an independent mutation in FLFW and the other three share a sweep.
*5. FLFW has an independent mutation and the others share a beneficial allele via standing variation

I've written a giant function to run all five models and calculate their composite likelihoods. The composite likelihoods include knowing how far away we are from the proposed selected site. Therefore, the first step for calculating them is to randomize with respect to the reference allele.  

The inverses and determinants generated can be used to calculate the log-likelihood of a site a given distance away from a proposed selected site. To do this, we sum the log-likelihoods across all SNPs in the window to obtain a composite log-likelihood under a given convergence model with a set of parameters for a proposed selected site. 

```{r runDMC,echo=TRUE,eval=FALSE}
```

I have to set some parameters before running the function, including the sets of parameters I want to try different settings for.

```{r setParameters}
```

And I'm going to run it with a variety of effective population sizes and recombination rates to see how robust the models are to that parameter.

```{r dmcNe, eval=FALSE}
```

I'm running this one on abies using `nohup Rscript ~/Projects/popgen/scripts/002_run_dmc.R >> ~/Projects/popgen/dmc_ne.log 2>&1 &`. 

## Estimate recombination rates

```{r rec_est,eval=FALSE}
ssc.map<-read.delim("../../scovelli_genome/SSC_linkage_map_20160701.txt",header=T)
colnames(ssc.map)<-c("map","mkr","gen","scaffold","scaffold_orientation","scaffold.bp","phys","node")
ssc.map$set<-"Syngnathus.scovelli"
map<-ssc.map[,c("set","map","mkr","phys","gen")]
write.table(map,"../../scovelli_genome/ssc_map.txt",sep=" ",row.names=FALSE, col.names = TRUE)
library(MareyMap)
startMareyMapGUI()
#using the GUI, I ran all three types of recombination rate estimators and saved the MapSet
```
```{r estimate_rec,eval=TRUE}
ssc.map<-read.delim("../../scovelli_genome/ssc_MareyMap.txt",comment.char = "#",sep=" ")
#genome-wide
r<-data.frame(Genome_wide=c(mean(ssc.map$slidingwindow,na.rm = TRUE)/1000000,
                            mean(ssc.map$spline,na.rm = TRUE)/1000000,
                            mean(ssc.map$loess,na.rm = TRUE)/1000000),
              LG4=c(mean(ssc.map$slidingwindow[ssc.map$map=="LG4"],na.rm = TRUE)/1000000,
                            mean(ssc.map$spline[ssc.map$map=="LG4"],na.rm = TRUE)/1000000,
                            mean(ssc.map$loess[ssc.map$map=="LG4"],na.rm = TRUE)/1000000),
              LG4_first12Mb=c(mean(ssc.map$slidingwindow[ssc.map$map=="LG4" & ssc.map$phys<12*10^6],na.rm = TRUE)/1000000,
                            mean(ssc.map$spline[ssc.map$map=="LG4"& ssc.map$phys<12*10^6],na.rm = TRUE)/1000000,
                            mean(ssc.map$loess[ssc.map$map=="LG4"& ssc.map$phys<12*10^6],na.rm = TRUE)/1000000),
              LG4_first10Mb=c(mean(ssc.map$slidingwindow[ssc.map$map=="LG4" & ssc.map$phys<10*10^6],na.rm = TRUE)/1000000,
                            mean(ssc.map$spline[ssc.map$map=="LG4"& ssc.map$phys<10*10^6],na.rm = TRUE)/1000000,
                            mean(ssc.map$loess[ssc.map$map=="LG4"& ssc.map$phys<10*10^6],na.rm = TRUE)/1000000),
              row.names = c("Sliding Window","Spline","Loess"))
kable(r,caption="Recombination rate estimates")
```
```{r plot_recLG4,eval=TRUE}
load("../../scovelli_genome/ssc_map.rda")
#plot LG4
plot(set[["LG4"]]@interpolations$slidingwindow@physicalPositions,
     set[["LG4"]]@interpolations$slidingwindow@rates,type="l",
     ylim=c(-5,25),xlab="",ylab="",xaxt='n',lwd=2)
abline(h=0,lty=2)
axis(1,at=seq(0,1.8*10^7,3*10^6),labels = seq(0,18,3))
mtext("Position on LG4 (Mb)",1,line=2)
mtext("Recombination Rate (cM/Mb)",2,line=2)
lines(set[["LG4"]]@interpolations$spline@physicalPositions,
      set[["LG4"]]@interpolations$spline@rates,type="l",col="blue",lwd=2)
lines(set[["LG4"]]@interpolations$loess@physicalPositions,
      set[["LG4"]]@interpolations$loess@rates,type="l",lwd=2,col="forestgreen")
legend("topleft",bty='n',col=c("black","blue","forestgreen"),lwd=2,c("Sliding Window","Spline","Loess"))
```


## Evaluate the output of the initial dmc runs

### Plot maximum composite likelihood ratios for models over proposed selected sites

```{r plot_likelihoods_fxns}
calc.max.complike<-function(pattern,neutral_name=NA){
  #read in composite likelihood files and calculate max for all proposed selected sites
  if(is.na(neutral_name)){
    compLikelihood_neutral = readRDS(paste("dmc/compLikelihood_neutral_",pattern,".RDS",sep=""))
  }else{
    compLikelihood_neutral = readRDS(neutral_name)
  }
  
  compLikelihood_neutral_site = sapply(1 : length(compLikelihood_neutral), function(i) {
    max(unlist(compLikelihood_neutral[[i]]))
  })
  
  compLikelihood_ind = readRDS(paste("dmc/compLikelihood_ind_",pattern,".RDS",sep=""))
  compLikelihood_ind_site = sapply(1 : length(compLikelihood_ind), function(i) {
    max(unlist(compLikelihood_ind[[i]]))
  })
  
  compLikelihood_mig = readRDS(paste("dmc/compLikelihood_mig_",pattern,".RDS",sep=""))
  compLikelihood_mig_site = sapply(1 : length(compLikelihood_mig), function(i) {
    max(unlist(compLikelihood_mig[[i]]))
  })
  
  compLikelihood_sv = readRDS(paste("dmc/compLikelihood_sv_",pattern,".RDS",sep=""))
  compLikelihood_sv_site = sapply(1 : length( compLikelihood_sv), function(i) {  
    max(unlist(compLikelihood_sv[[i]]))
  })
  
  compLikelihood_mixed_migInd = readRDS(paste("dmc/compLikelihood_mixed_migInd_",pattern,".RDS",sep=""))
  compLikelihood_mixed_migInd_site = sapply(1 : length(compLikelihood_mixed_migInd), function(i) {
    max(unlist(compLikelihood_mixed_migInd[[i]]))
  })
  
  compLikelihood_mixed_svInd = readRDS(paste("dmc/compLikelihood_mixed_svInd_",pattern,".RDS",sep=""))
  compLikelihood_mixed_svInd_site = sapply(1 : length(compLikelihood_mixed_svInd), function(i) {
    max(unlist(compLikelihood_mixed_svInd[[i]]))
  })
  return(list(max.likes=c(ind=(compLikelihood_ind_site - compLikelihood_neutral_site),
           mig=(compLikelihood_mig_site - compLikelihood_neutral_site), 
           sv=(compLikelihood_sv_site - compLikelihood_neutral_site),
           migInd=(compLikelihood_mixed_migInd_site - compLikelihood_neutral_site),
           svInd=(compLikelihood_mixed_svInd_site - compLikelihood_neutral_site)),
         max.complikes=c(neutral=compLikelihood_neutral,ind=compLikelihood_ind,
           mig=compLikelihood_mig, 
           sv=compLikelihood_sv,
           migInd=compLikelihood_mixed_migInd,
           svInd=compLikelihood_mixed_svInd)))
}

plot.complike<-function(pattern,selSite,leg=TRUE,lab=TRUE){
  max.complikes<-calc.max.complike(pattern)
  max.likes<-max.complikes[[1]]
  plot_range = range(max.likes)
  
  if(lab==TRUE){
    x.lab<-"Proposed position selected site"
    y.lab<-"Composite log-likelihood (model - neutral)"
  } else{
    x.lab<-y.lab<-""
  }
  
  plot(selSite, max.likes[grep("ind",names(max.likes))], type = "b",
       ylim = c(plot_range[1] - 50, plot_range[2] + 50),
       xlab = x.lab,
       ylab = y.lab)
  lines(selSite, max.likes[grep("mig\\d+",names(max.likes))], col = "red",
        type = "b")
  lines(selSite, max.likes[grep("sv\\d+",names(max.likes))], col = "blue",
        type = "b")
  lines(selSite, max.likes[grep("migInd",names(max.likes))],
        col = "orange", lty = 2, type = "b")
  lines(selSite, max.likes[grep("svInd",names(max.likes))],
        col = "green", lty = 2, type = "b")
  legend("top",bty='n',legend = pattern)
  if(isTRUE(leg)){
    legend("topright", col = c("black", "red", "blue", "orange", "green"),
           lty = c(rep(1, 3), rep(2, 2)), sapply(1 : 5, function(i) paste("Model", i)))
  }
  return(list(max.complikes[[2]],max.likes,pattern))
}

```

```{r plot_ne,fig.width=8,fig.height=8}
patterns<-c("p4LG4_8300_2_8","p4LG4_83000_2_8","p4LG4_830000_2_8","p4LG4_8300000_2_8",
            "p4LG4_8300_2_7","p4LG4_83000_2_7","p4LG4_830000_2_7","p4LG4_8300000_2_7",
            "p4LG4_8300_6_6","p4LG4_83000_6_6","p4LG4_830000_6_6","p4LG4_8300000_6_6",
            "p4LG4_8300_5_6","p4LG4_83000_5_6","p4LG4_830000_5_6","p4LG4_8300000_5_6")
selSite = positions[seq(1, length(positions), length.out = 50)]
par(mfrow=c(4,4),mar=c(1.5,1.5,0.5,0.5),oma=c(2,2,2,2))
ne.complikes<-lapply(patterns,plot.complike,leg=FALSE,lab=FALSE,selSite=selSite)
mtext("Position on LG4",1,outer=TRUE,cex=0.8,line=.5)
mtext("Composite log likelihood (model - neutral)",2,outer=T,cex=0.8,line=.5)
par(fig=c(0,1,0,1),oma=c(0,0,0,0),mar=c(0,0,0,0),new=TRUE)
plot(0,0,type='n',bty='n',xaxt='n',yaxt='n')
legend("top", col = c("black", "red", "blue", "orange", "green"),ncol=5,bty='n',
           lty = c(rep(1, 3), rep(2, 2)), sapply(1 : 5, function(i) paste("Model", i)))
```

### Get maximum composite likelihood estimates

```{r get_mcle_fxn}
#These won't work if the parameters aren't specified as above
get.mcle<-function(max.complikes,Ne,rec,selSite= positions[seq(1, length(positions), length.out = 50)],
                   sels = c(1e-4, 1e-3, 0.01, seq(0.02, 0.14, by = 0.01), seq(0.15, 0.3, by = 0.05), 
                            seq(0.4, 0.6, by = 0.1)),
                   times = c(0, 5, 25, 50, 100, 500, 1000, 1e4, 1e6),
                   migs = c(10^-(seq(5, 1, by = -2)), 0.5, 1),sources=c(3,5,7,16)){
  Ne<<-Ne
  rec<<-rec
  gs = c(1/(2*Ne), 10^-(4:1))
  source("../programs/dmc-master/getMCLE.R")

  ## Model 1
  m1<-cbind(getMCLEind(max.complikes[grep("ind",names(max.complikes))], selSite, sels),
            maxG=NA,maxTime=NA,maxSource=NA)
  
  ## Model 2
  m2<-cbind(getMCLEmig(max.complikes[grep("mig\\d+",names(max.complikes))], selSite, sels, migs, sources),
            maxG=NA)

  ## Model 3
  m3<-getMCLEsv_source(max.complikes[grep("sv\\d+",names(max.complikes))], selSite, sels, gs, times, sources)
  
  ## Model 4
  m4<-getMCLEmixed(max.complikes[grep("migInd",names(max.complikes))], selSite, sels, gs[1], times[1], migs, sources)
  
  ## Model 5
  m5<-getMCLEmixed(max.complikes[grep("svInd",names(max.complikes))], selSite, sels, gs, times, migs[1], sources)
  mcles<-rbind(m1,m2,m3,m4,m5)
  rownames(mcles)<-c("Model1","Model2","Model3","Model4","Model5")
  return(mcles)
}
```

```{r get_mcle}
rs<-c("2_8"=2.17*10^-8,"2_7"=2*10^-7,"6_6"=6*10^-6,"5_6"=5.6*10^-6)
ne.mcles<-lapply(ne.complikes, function(m.complikes){
  pattern<-m.complikes[[3]]
  print(pattern)
  Ne<-as.numeric(gsub("p4LG4_(\\d+)_.*","\\1",pattern))
  r<-gsub("p4LG4_\\d+_(\\d_\\d)","\\1",pattern)
  rec<-rs[r]
  max.comp<-m.complikes[[1]]
  mcle<-get.mcle(max.comp,Ne=Ne,rec=rec)
  return(list(mcle,Ne,rec))
})
names(ne.mcles)<-patterns
lapply(ne.mcles,function(mcle){ kable(mcle[[1]]) })
```

## Model with chosen parameters

```{r dmcTwoNes, eval=FALSE}
```

```{r plot_chosen,eval=TRUE,fig.height=10,fig.width=6,fig.align='H'}
selSite = positions[seq(1, length(positions[positions<12000000]), length.out = 100)]
par(mfrow=c(2,1),oma=c(2,2,2,2),mar=c(2,2,2,2))
max.complikes1<-plot.complike("p4LG4_100000_2_8",leg=FALSE,selSite = selSite)
max.complikes2<-plot.complike("p4LG4_1000000_2_8",leg=FALSE,selSite = selSite)
mcle1<-get.mcle(max.complikes1[[1]],Ne=100000,rec=2*10^-8,selSite = selSite)
kable(mcle1)
mcle2<-get.mcle(max.complikes2[[1]],Ne=1000000,rec=2*10^-8,selSite = selSite)
kable(mcle2)
```

The selection values are still the smallest ones. Kristin Lee indicated that this might occur if there aren't enough neutral SNPs - so I'll try re-calculating neutral SNPS

```{r recalc_neutral,eval=FALSE}
#calculate selected allele frequencies
LG4.sub<-calc.allFreqs(vcf[vcf$`#CHROM` == "LG4" & vcf$POS < 12*10^6,],pop.list,pop.labs)
saveRDS(LG4.sub,"dmc/selectedRegionAlleleFreqs_p4LG4_subset.RDS")
saveRDS(vcf[vcf$`#CHROM`=="LG4"& vcf$POS < 12*10^6,"POS"],"dmc/selectedRegionPositions_p4LG4_subset.RDS")

#calculate neutral allele frequencies
all.out<-read.delim("all_outliers.txt",sep='\t',header=TRUE)
allFreqs<-calc.allFreqs(vcf[!vcf$`#CHROM` %in% "LG4" & !vcf$SNP %in% all.out$SNP,],
                   pop.list,pop.labs)
saveRDS(allFreqs,"dmc/neutralAlleleFreqs_p4LG4_notSNPs.RDS")

neutralF_filename<-"dmc/neutralF_p4LG4_subset"
source("../programs/dmc-master/calcNeutralF.R")
F_estimate<-readRDS("dmc/neutralF_p4LG4_subset.RDS")

#calculate the determinant and inverse of the neutral F matrix
M <- numPops
Tmatrix <- matrix(data = rep(-1 / M, (M - 1) * M), nrow = M - 1, ncol = M)
diag(Tmatrix) = (M - 1) / M 
sampleErrorMatrix = diag(1/sampleSizes, nrow = numPops, ncol = numPops)
det_FOmegas_neutral = det(Tmatrix %*% (F_estimate + sampleErrorMatrix) %*% t(Tmatrix))
saveRDS(det_FOmegas_neutral, "dmc/det_FOmegas_neutral_p4LG4_subset.RDS")

inv_FOmegas_neutral = ginv(Tmatrix %*% (F_estimate + sampleErrorMatrix) %*% t(Tmatrix))
saveRDS(inv_FOmegas_neutral, "dmc/inv_FOmegas_neutral_p4LG4_subset.RDS")

```
```{r dmcSubset, eval=FALSE}
```
```{r plot_subset,fig.height=10,fig.width=10,fig.align='H'}
positions<-readRDS("dmc/selectedRegionPositions_p4LG4_subset.RDS")
F_estimate<-readRDS("dmc/neutralF_p4LG4_subset.RDS")
sampleSizes<-readRDS("dmc/sampleSizes.RDS")
neutral_det_name = "dmc/det_FOmegas_neutral_p4LG4_subset.RDS"
neutral_inv_name = "dmc/inv_FOmegas_neutral_p4LG4_subset.RDS"
selSite = positions[seq(1, length(positions[positions<12000000]), length.out = 100)]
par(mfrow=c(1,1),oma=c(2,2,2,2),mar=c(2,2,2,2))
max.complikes1<-plot.complike("p4LG4_100000_2_8_sub",leg=TRUE,selSite = selSite)
mcle.sub<-get.mcle(max.complikes1[[1]],Ne=100000,rec=2*10^-8,
                   selSite=selSite)
kable(mcle.sub)
```

### Let's try lower recombination rates

```{r dmcRecomb, eval=FALSE}
```

```{r plot_recomb,eval=TRUE,fig.height=10,fig.width=6,fig.align='H'}
selSite = positions[seq(1, length(positions), length.out = 50)]
F_estimate<-readRDS("dmc/neutralF_p4LG4_50.RDS")
sampleSizes<-readRDS("dmc/sampleSizes.RDS")
par(mfrow=c(2,1),oma=c(2,2,2,2),mar=c(2,2,2,2))
max.complikes1<-plot.complike("p4LG4_100000_2_9",leg=FALSE,selSite = selSite)
max.complikes2<-plot.complike("p4LG4_100000_2_10",leg=FALSE,selSite = selSite)
mcle1<-get.mcle(max.complikes1[[1]],Ne=100000,rec=2*10^-9)
kable(mcle1)
mcle2<-get.mcle(max.complikes2[[1]],Ne=100000,rec=2*10^-10)
kable(mcle2)
```

### Different selection parameters

*Going below 1e-5 (i.e. to 1e-6) caused infinite values to occur*

```{r dmcSelection,eval=FALSE}
```

```{r plot_sel,eval=TRUE,fig.height=6,fig.width=6,fig.align='H'}
sels = c(9e-4,7.5e-4,5e-4, 2.5e-4,1e-4)
selSite = positions[seq(1, length(positions[positions<12000000]), length.out = 50)]
F_estimate<-readRDS("dmc/neutralF_p4LG4_50.RDS")
par(mfrow=c(1,1),oma=c(2,2,2,2),mar=c(2,2,2,2))
max.complikes.sel<-plot.complike("p4LG4_100000_2_9_sels",leg=FALSE,selSite = selSite)
mcle.sel<-get.mcle(max.complikes.sel[[1]],Ne=100000,rec=2*10^-9,sels=sels)
kable(mcle.sel)

```

### Chosen parameter runs

Using recombination rate of 2e-9, Ne= 100000, with a wider range of selection values that go below 1e-4, and 100 selection sites from 0 to 10Mb.

```{r dmcForReal,eval=FALSE}
```

```{r plot_final,eval=TRUE,fig.height=6,fig.width=6,fig.align='H'}
selSite = positions[seq(1, length(positions[positions<10000000]), length.out = 100)]
F_estimate<-readRDS("dmc/neutralF_p4LG4_100_10MB.RDS")
    sels = c(1e-4, 1e-3, 0.01, seq(0.02, 0.14, by = 0.01), seq(0.15, 0.3, by = 0.05), 
             seq(0.4, 0.6, by = 0.1))
    par(mfrow=c(1,1),oma=c(2,2,2,2),mar=c(2,2,2,2))
max.complikes<-plot.complike("p4LG4_100000_2_9_forReal",leg=FALSE,selSite = selSite)
mcle<-get.mcle(max.complikes[[1]],Ne=100000,rec=2*10^-9,sels=sels,selSite=selSite)
kable(mcle)

```

## Plot profile likelihood surfaces

Plot the profile likelihood surfaces for Models 3 and 5, where standing variation is present in all or one population, which have the highest composite log-likelihoods. 

```{r plot_likeProfile,eval=TRUE,fig.height=10,fig.width=6,fig.align='H'}
selSite = positions[seq(1, length(positions[positions<10000000]), length.out = 100)]
gs<-c(1/(2*Ne), 10^-(4:1))
Ne=100000
rec=2*10^-9
sources <- c(3,5,7,16)
times = c(0, 5, 25, 50, 100, 500, 1000, 1e4, 1e6)
migs = c(10^-(seq(5, 1, by = -2)), 0.5, 1)
    sels = c(1e-4, 1e-3, 0.01, seq(0.02, 0.14, by = 0.01), seq(0.15, 0.3, by = 0.05), 
             seq(0.4, 0.6, by = 0.1))
par(mfrow=c(2,1),oma=c(2,2,2,2),mar=c(2,2,2,2))
## Model 3
compLikelihood_sv<-readRDS("dmc/compLikelihood_sv_p4LG4_100000_2_9_forReal.RDS")
names(compLikelihood_sv)<-selSite
compLikelihood_sv_site = sapply(1 : length(selSite), function(i) {  
  max(unlist(compLikelihood_sv[[i]]))
})
mcle_sv = getMCLEsv_source(compLikelihood_sv,selSite, sels, gs, times, sources)
mcle_sv_index<-which(mcle_sv[1]==selSite)
compLike_sv_byTime = lapply(1 : length(times), function(time) {
  sapply(1: length(sels), function(sel) {
    sapply(1 : length(gs), function(g) {
      compLikelihood_sv[[mcle_sv_index]][[sel]][[g]][[time]]
    })
  })
})

profileLike_time_sv = sapply(1: length(compLike_sv_byTime), function(i) {
  max(unlist(compLike_sv_byTime[[i]]))
})

plot(times, profileLike_time_sv, type = "b", xlab = "Time",
     ylab = "Profile composite log-likelihood", main = "Figure 2: Model 3")
abline(v = mcle_sv[4], lty = 2, col = "red")

## Model 5
compLikelihood_mixed_svInd<-readRDS("dmc/compLikelihood_mixed_svInd_p4LG4_100000_2_9_forReal.RDS")
mcle_mixed_svInd = getMCLEmixed(compLikelihood_mixed_svInd, selSite, sels,
                                gs, times, migs[1], sources)
mcle_mixed_svInd_index<-which(mcle_sv[1]==selSite)
compLike_mixed_svInd_byTime = lapply(1 : length(times), function(time) {
  sapply(1: length(sels), function(sel) {
    sapply(1 : length(gs), function(g) {
      compLikelihood_mixed_svInd[[mcle_mixed_svInd_index]][[sel]][[g]][[time]]
    })
  })
})

profileLike_time_mixed_svInd = sapply(1: length(compLike_mixed_svInd_byTime), function(i) {
  max(unlist(compLike_mixed_svInd_byTime[[i]]))
})

plot(times, profileLike_time_mixed_svInd, type = "b", xlab = "Time",
     ylab = "Profile composite log-likelihood", main = "Figure 3: Model 5")
abline(v = mcle_mixed_svInd[4], lty = 2, col = "red")
```


We see little difference in the composite log-likelihoods of models 3 and 5, provides further evidence that independent mutations (or selection very old standing variation) in FLFW generated the patterns observed in the data. The point estimates of MCLE of t are marked by red lines in Figures 2 and 3.

## MCLE for the strength of selection

The best fitting models are 5 (green), 3 (blue), and 1 (black). Models 3 and 5 estimate selection at 0.2 and model 1 estimates it at 0.15. 

```{r mcle_s}
## Model 1
compLikelihood_ind<-readRDS("dmc/compLikelihood_ind_p4LG4_100000_2_9_forReal.RDS")
mcle_ind = getMCLEind(compLikelihood_ind, selSite, sels)
mcle_ind_index<-which(mcle_ind[1]==selSite)
profileLike_sel_ind = sapply(1: length(sels), function(i) {
  max(unlist(compLikelihood_ind[[mcle_ind_index]][[i]]))
})

## Model 3
profileLike_sel_sv = sapply(1: length(sels), function(i) {
  max(unlist(compLikelihood_sv[[mcle_sv_index]][[i]]))
})

## Model 5
profileLike_sel_mixed_svInd = sapply(1: length(sels), function(i) {
  max(unlist(compLikelihood_mixed_svInd[[mcle_mixed_svInd_index]][[i]]))
})

par(mfrow = c(1, 3))
plot(sels, profileLike_sel_sv, type = "b", xlab = "Sel",
     ylab = "Profile composite log-likelihood", main = "Model 3")
abline(v = mcle_sv[2], lty = 2, col = "blue")

plot(sels, profileLike_sel_mixed_svInd, type = "b", xlab = "Sel",
     ylab = "Profile composite log-likelihood", main = "Model 5")
abline(v = mcle_mixed_svInd[2], lty = 2, col = "blue")

plot(sels, profileLike_sel_ind, type = "b", xlab = "Sel",
     ylab = "Profile composite log-likelihood", main = "Model 1")
abline(v = mcle_ind[2], lty = 2, col = "blue")

```

## Where is the selected site?

The selected site is somewhere between $`r min(mcle_ind[1],mcle_sv[1],mcle_mixed_svInd[1])`$ and $`r max(mcle_ind[1],mcle_sv[1],mcle_mixed_svInd[1])`$, if we take the maximum estimates from models 1, 3, and 5.

```{r plot_final_again}
par(mfrow=c(1,1),oma=c(2,2,2,2),mar=c(2,2,2,2))
max.complikes<-plot.complike("p4LG4_100000_2_9_forReal",leg=FALSE,selSite = selSite)
abline(v=mcle_ind[1],lty=2,lwd=2,col="black")
abline(v=mcle_sv[1],lty=2,lwd=2,col="blue")
abline(v=mcle_mixed_svInd[1],lty=2,lwd=2,col="green")
```

