---
title: "FWSW analysis with dadi"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.extra='',fig.pos="H")
knitr::opts_knit$set(root.dir='./fwsw_results/')
```
```{r source}
source("../../gwscaR/R/gwscaR.R")
source("../../gwscaR/R/gwscaR_plot.R")
source("../../gwscaR/R/gwscaR_utility.R")
source("../../gwscaR/R/gwscaR_fsts.R")
source("../../gwscaR/R/gwscaR_popgen.R")
source("../../gwscaR/R/vcf2dadi.R")
library(knitr)
pop.list<-c("ALFW","ALST","FLCC","FLLG","LAFW","TXCC","TXFW")
```



# Create a dadi SNPs file from the vcf

```{r create_dadi}
dadi<-vcf2dadiSNPs(vcf,pop.list = pop.list,filename = "dadi_analysis/fwsw.dadi.snps")
dadi<-read.delim("dadi_analysis/fwsw.dadi.snps")
projections<-unlist(lapply(pop.list,function(pop){ n<-length(grep(pop,colnames(vcf)))}))*2

names(projections)<-pop.list
best.pops<-c("TXFW","TXCB","LAFW","ALST","ALFW","FLLG","FLCC","TXFW ","TXCB ","LAFW ","ALST ","ALFW ","FLLG ","FLCC ")
best.pops<-c("TXFW","TXCB","LAFW","ALST","ALFW","FLLG","FLCC","TXFW.1","TXCB.1","LAFW.1","ALST.1","ALFW.1","FLLG.1","FLCC.1")
dadi<-dadi[,c("Ingroup","Outgroup","Allele1",best.pops[1:7],"Allele2",best.pops[8:14],"GeneID","Position")]
dim(dadi[rowSums(dadi[,colnames(dadi) %in% best.pops])==sum(projections[best.pops[1:7]]),])

#transform this and remove everything but population info for ease of operations
dd<-as.data.frame(cbind(colnames(dadi)[colnames(dadi) %in% best.pops],t(dadi[,colnames(dadi) %in% best.pops])),row.names = 0,stringsAsFactors = FALSE)
#rename the pops to be factors, essentially
dd$V1<-gsub("\\.1","",dd$V1)
#calculate sums per locus for each population
ns<-data.frame(do.call(rbind,lapply(2:ncol(dd),function(c){ tapply(dd[,c],dd$V1,function(x){ sum(as.numeric(x)) }) })))
rownames(ns)<-dadi$GeneID
nozeros<-ns[!(apply(ns, 1, function(y) any(y == 0))),] #512 had zeroes
props<-apply(nozeros,2,function(x) { p<-x/max(x) } )
keepers<-props[apply(props, 1, function(y) any(y >=0.75)),]

#now save these in the dadi format
write.table(dadi[dadi$GeneID %in% rownames(keepers),],"dadi_analysis/fwsw.dadi.pruned.snps",quote=FALSE,row.names = FALSE,
            col.names = c("Ingroup","Outgroup","Allele1","TXFW","TXCB","LAFW","ALST","ALFW","FLLG","FLCC","Allele2","TXFW","TXCB","LAFW","ALST","ALFW","FLLG","FLCC","GeneID","Position"))
```

```{r check_dadis}
vcf.files<-c("stacks/populations_subset50/batch_2.vcf",
             "stacks/populations_subset75/batch_2.vcf",
             #"stacks/populations_subset85/batch_2.vcf",
             #"stacks/populations_subset90/batch_2.vcf",
             #"stacks/populations_subset95/batch_2.vcf",
             "stacks/populations_subset100/batch_2.vcf")

dadis<-lapply(vcf.files,function(file){
  full.vcf<-parse.vcf(file)
  vcf<-choose.one.snp(full.vcf)
  vcf.out.name<-gsub("batch_2.vcf","batch_2.pruned.vcf",file)
  write.table(x = vcf,file = vcf.out.name,quote = FALSE,sep='\t',row.names = FALSE,col.names = TRUE)
  dadi.out<-paste("dadi_analysis/fwsw",gsub(".*populations_subset(\\d+).*","\\1",file),".dadi.snps",sep="")
  pop.list<-c("ALFW","ALST","FLCC","FLLG","LAFW","TXCC","TXFW")
  dadi<-vcf2dadiSNPs(vcf,pop.list = pop.list,filename = dadi.out)
  return(dadi)
})

#compare
kable(cbind(file=vcf.files,numer_snps=unlist(lapply(dadis,nrow))))
```

I'm going to use th 75% of individuals dataset.

## Analyze 1D dadi results

I ran the optimize functions from (dportik's dadi pipeline)[https://github.com/dportik/dadi_pipeline] for each of the 7 populations. Now I need to see what the optimal parameters are (with lowest AIC).

```{r}
opt.params<-do.call(rbind,lapply(pop.list, function(pop,dadi.dir="dadi_analysis"){
  if(length(grep("/^",dadi.dir))==0){
    dadi.dir<-paste(dadi.dir,"/",sep="")
  }
  path<-paste(dadi.dir,pop,sep="")
  opt.files<-list.files(path = path,pattern = "optimized",full.names = TRUE)
  opts<-do.call(rbind,lapply(opt.files,function(opt.file){
    model.opt<-read.delim(opt.file,sep='\t',header=TRUE)
    model.opt<-model.opt[-which(model.opt$Model=="Model"),]
    opt<-model.opt[which.min(as.numeric(as.character(model.opt$AIC))),]
    return(opt)
  }))
  opts$pop<-pop
  return(opts)
}))
```

Once I have that, then I'll run the optimal set and do model comparison to choose the most likely single-population scenarios. How do I choose the best model? Maybe use the dadi imaging functions to look at model fits. Also, likelihood ratio tests could be useful. 









